---
title: "Mapping with r"
author: "Lena Moser"
date: "2023-01-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Week 48

3) Create Interactive Maps in R: Mapping places in Aarhus

```{r}
# Activate the libraries
library(leaflet)
library(htmlwidgets)
```

Task 1: Create a Danish equivalent of AUSmap with esri layers but called DANmap

```{r task 1, echo=TRUE}
l_dan <- leaflet() %>%   # assign the base location to an object
  setView(10.2033485, 56.1724346, zoom = 13) # coordinater must be opposite to Aus, since we are on the opposite half of globe

# Now, prepare to select backgrounds - selects all addtiles starting with esri, finding all esri maps available without
# typing individual maps
esri <- grep("^Esri", providers, value = TRUE)

# Select backgrounds from among provider tiles. To view them the options, 
# go to https://leaflet-extras.github.io/leaflet-providers/preview/
for (provider in esri) {
  l_dan <- l_dan %>% addProviderTiles(provider, group = provider)
}

# run object to see map - we have created a basemap
l_dan


# We make a layered map out of the components above and write it to 
# an object called DANmap
DANmap <- l_dan %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")

# run this to see your product
DANmap
```

########## ADD DATA TO LEAFLET

In this section you will manually create machine-readable spatial
data from GoogleMaps: 

First, go to https://bit.ly/CreateCoordinates1
Enter the coordinates of your favorite leisure places in Denmark 
      # extracting them from the URL in googlemaps, adding name and type of monument.
      # Remember to copy the coordinates as a string, as just two decimal numbers separated by comma. 

Caveats: Do NOT edit the grey columns! They populate automatically!

Second, read the sheet into R. You will need gmail login information. 
      # watch the console, it may ask you to authenticate or put in the number 
      # that corresponds to the account you wish to use.


```{r, echo=TRUE}
# Libraries
library(tidyverse)
library(googlesheets4)
library(leaflet)

gs4_deauth()

# Read in a Google sheet
places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
                     col_types = "cccnncnc", range = "SA2022")


glimpse(places)

# load the coordinates in the map and check: are any points missing? Why?
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Placename)

```
```{r task 2, echo=TRUE}
# Task 2: Read in the googlesheet data you and your colleagues 
# populated with data into the DANmap object you created in Task 1.
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description)

```

```{r task 3, echo=TRUE}
# Task 3: Can you cluster the points in Leaflet? Google "clustering options in Leaflet"
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description,
             clusterOptions = markerClusterOptions())
```

Task 4: Look at the map and consider what it is good for and what not.
It shows areas of interest and how they cluster on the map, but little information is available

```{r task 5, echo=TRUE}
# Task 5: Find out how to display notes and classifications in the map.
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             label = places$Description,
             clusterOptions = markerClusterOptions(),
             popup = paste("<br>Type:", places$Type,
                           "<br>Notes:", places$Notes))
```

